---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
setInternet2(use = TRUE)
tmp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",tmp)
data <- read.csv(unz(tmp, "activity.csv"))
# convert the string date  to Date format
data$Date <- as.Date(data$date, "%Y-%m-%d")
unlink(tmp)
```


## What is mean total number of steps taken per day?
```{r}
dataagg <- aggregate(data$steps, by=list(Category=data$date), FUN=sum,na.rm=TRUE)
hist(dataagg$x, main="Steps per day",col="red",xlab="")

```

The mean total number of steps per day is `r sprintf("%.2f", mean(dataagg$x,na.rm=TRUE))` and the median number of steps per day is `r sprintf("%.2f",median(dataagg$x,na.rm=TRUE))`.

## What is the average daily activity pattern?

```{r}
aggint <- aggregate(data$steps, by=list(Category=data$interval), FUN=mean,na.rm=TRUE)
aggint <- aggregate(steps ~ interval,data = data,FUN=function(x) c(steps =mean(x,na.rm=TRUE) ))
plot(steps ~ interval, aggint, type = "l")

```

The 5-minute interval at `r aggint$Category[aggint$x == max(aggint$x)]`, on average across all the days in the dataset, contains the maximum number of steps (`r  max(aggint$x)`).

## Imputing missing values

```{r}
imputed <- data.frame(data)
repfunc <- function(a,b,c) 
  { if(is.na(a)) 
       c$steps[c$interval == b]
    else
      a
  }
for(row in 1:length(imputed$steps)) { 
    imputed$steps[row] <- repfunc(imputed$steps[row], imputed$interval[row],aggint) 
}
imputedagg <- aggregate(imputed$steps, by=list(Category=imputed$date), FUN=sum,na.rm=TRUE)
imputedagg <- aggregate(steps ~ date,data = imputed,FUN=function(x) c(steps =sum(x) ) )

hist(imputedagg$steps, main="Steps per day",col="red",xlab="")

meandiff <- mean(imputedagg$steps,na.rm=TRUE) - mean(dataagg$x,na.rm=TRUE)
             
if (mean(imputedagg$steps,na.rm=TRUE) > mean(dataagg$x,na.rm=TRUE))
  difftext <- "more" else  difftext <- "less"

```

The number of rows with missing data is `r sum(is.na(data$steps))`.
The mean total number of steps per day is `r sprintf("%.2f", mean(imputedagg$x,na.rm=TRUE))` and the median number of steps per day is `r sprintf("%.2f",median(imputedagg$x,na.rm=TRUE))`.

There are `r meandiff ` `r difftext ` mean daily steps in the imputed  data than in the original data. 

## Are there differences in activity patterns between weekdays and weekends?


```{r}

repday <- function(a) 
  { c("weekend", "weekday", "weekday", "weekday", "weekday", 
    "weekday", "weekend")[as.POSIXlt(a)$wday + 1]
  }
for(row in 1:length(imputed$steps)) { 
    imputed$day[row] <- repday(imputed$date[row]) 
}
imputed$day <- as.factor(imputed$day) 
#  panel plot containing a time series plot (i.e. type = "l") of the 5-min
library("lattice")
aggday <- aggregate(. ~ interval+day,data = imputed,FUN=function(x) c(steps =mean(x) ) )
xyplot(steps ~ interval | day,
        data = aggday,
        type = "l",
        layout = c(1,2))

```
